\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2022-04-25 sections/methods.tex
Add: Add captions and update supplementary figures

  \subsection*{Denoising and Clustering (DC)}
  \vspace{-5mm}
  This module deals with processing the quality-controlled, trimmed 16S sequence data into \ac{otu} or \ac{esv} count tables.
  It consists of the following processes: denoising (or clustering) and chimera checking.
  The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables (some methods, like closed reference and open reference clustering, perform clustering and taxonomy assignment in the same step).
  The chimera checking process handles the removal of chimeric sequences created during the \ac{pcr} step.
  The output of this module is a matrix of counts, that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix).
  The options currently available in the pipeline for denoising and clustering are: open reference clustering, closed reference clustering and de novo clustering methods from the vsearch plugin of \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019} and denoising methods from \ac{dada2} v1.14~\cite{Callahan2016} (from the \ac{dada2} R package) and Deblur v1.1.0~\cite{Amir2017} (from the deblur plugin of \ac{qiime2}).
  The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} v2021.8.0 (uchime method) and \ac{dada2} (remove bimera method).
  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:dc_tools}.

@@

  \begin{equation*}
    g(x) =
    \begin{cases}
       & 0, \text{ if } x=0, \\
       & -1, \text{ if } x<0, \\
       & 1, \text{ if } x>0
    \end{cases}
  \end{equation*}
  and
  \begin{equation*}
    f(x_1,\dots,x_w) = max \left( \#(i \mid x_i=-1),\#(i \mid x_i=1)
       \right)
  \end{equation*}
  where, $#$ refers to the cardinality of the set.

@@

  \begin{equation*}
   \mathbf{\bar{N}} =
     \begin{blockarray}{*{4}{c} l}
      \begin{block}{*{4}{>{$\footnotesize}c<{$}} l}
        $\bar{N}$^1 & $\bar{N}$^2 & \dots & $\bar{N}$^{W} & \\
      \end{block}
      \begin{block}{[*{4}{c}]>{$\footnotesize}l<{$}}
       n^1_1 & n^2_1 & \cdots  & n^{W}_1 \bigstrut[t]& $\bar{N}$_1 \\
       n^1_2 & n^2_2 & \cdots  & n^{W}_2 & $\bar{N}$_2 \\
       \vdots & \vdots & \vdots  & \vdots & \vdots \\
       n^1_{Q^2} & n^2_{Q^2} & \cdots  & n^{W}_{Q^2} & $\bar{N}$_{Q^2}
      \end{block}
     \end{blockarray}
  \end{equation*}

>>>>

  \subsection*{Denoising and Clustering (DC)}
  \vspace{-5mm}
  This module deals with processing the quality-controlled, trimmed 16S sequence data into \ac{otu} or \ac{esv} count tables.
  It consists of the following processes: denoising (or clustering) and chimera checking.
  The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables (some methods, like closed reference and open reference clustering, perform clustering and taxonomy assignment in the same step).
  The chimera checking process handles the removal of chimeric sequences created during the \ac{pcr} step.
  The output of this module is a matrix of counts, that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix).
  The options currently available in the pipeline for denoising and clustering are: open reference clustering, closed reference clustering and de novo clustering methods from the vsearch plugin of \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019} and denoising methods from \ac{dada2} v1.14~\cite{Callahan2016} (from the \ac{dada2} R package) and Deblur v1.1.0~\cite{Amir2017} (from the deblur plugin of \ac{qiime2}).
  The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} v2021.8.0 (uchime-denovo method) and \ac{dada2} (remove bimera method).
  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:dc_tools}.

@@

  \begin{equation*}
    g(x) =
    \begin{cases}
       & 0, \text{ if } x=0, \\
       & -1, \text{ if } x<0, \\
       & 1, \text{ if } x>0
    \end{cases}
  \end{equation*}
  and
  \begin{equation*}
    f(x_1,\dots,x_w) = max \left( \#(i \mid x_i=-1),\#(i \mid x_i=1)
       \right)
  \end{equation*}
  where, # refers to the cardinality of the set.

@@

  \begin{equation*}
   \mathbf{\bar{N}} =
     \begin{blockarray}{*{4}{c} l}
      \begin{block}{*{4}{>{$\footnotesize}c<{$}} l}
        \bar{N}^1 & \bar{N}^2 & \dots & \bar{N}^{W} & \\
      \end{block}
      \begin{block}{[*{4}{c}]>{$\footnotesize}l<{$}}
       n^1_1 & n^2_1 & \cdots  & n^{W}_1 \bigstrut[t]& \bar{N}_1 \\
       n^1_2 & n^2_2 & \cdots  & n^{W}_2 & \bar{N}_2 \\
       \vdots & \vdots & \vdots  & \vdots & \vdots \\
       n^1_{Q^2} & n^2_{Q^2} & \cdots  & n^{W}_{Q^2} & \bar{N}_{Q^2}
      \end{block}
     \end{blockarray}
  \end{equation*}

### 2022-04-25 sections/results.tex
Add: Add captions and update supplementary figures

  The recommended tool for the \ac{dc} step (\ac{dada2}) was chosen based on its accuracy in recapitulating the reference sequences in mock communities.
  For the chimera checking module of the \ac{dc} step, both uchime and remove bimera perform similarly, we suggest using remove bimera since it was developed specifically for \ac{dada2} and should provide the best performance for that denoising method.
  The choice of the taxonomy reference database in the \ac{ta} step is dictated largely by the species expected to be present in the sample as well the database used in similar studies, if comparison is a goal.
  Nevertheless, we suggest \ac{gg} database along with the NaiveBayes classifier as the query tool since the database is used most often in taxonomic studies.
  % TODO: Default OP step values and references
  The abundance threshold at the \ac{op} step is set at ... because large number of nodes are added to networks and decrease the performance of network inference algorithms
  For the \ac{op} step, we recommend that the filtering be turned on and used at the default settings, unless a taxonomy entity of interest might be filtered out at the default settings.
  This is necessary since a large number of rows in the counts matrix would reduce the accuracy of the inference methods and increase the computational time.
  Finally for the \ac{ni}, we recommend the scaled-sum consensus method along with p-value merging (for the correlation-based methods) to generate a final consensus network for the input data.

>>>>

  The recommended tool for the \ac{dc} step (\ac{dada2}) was chosen based on its accuracy in recapitulating the reference sequences in mock communities.
  For the chimera checking module of the \ac{dc} step, both uchime-denovo and remove bimera perform similarly, we suggest using remove bimera since it was developed specifically for \ac{dada2} and should provide the best performance for that denoising method.
  The choice of the taxonomy reference database in the \ac{ta} step is dictated largely by the species expected to be present in the sample as well the database used in similar studies, if comparison is a goal.
  Nevertheless, we suggest \ac{gg} database along with the NaiveBayes classifier as the query tool since the database is used most often in taxonomic studies.
  % TODO: Default OP step values and references
  The abundance threshold at the \ac{op} step is set at ... because large number of nodes are added to networks and decrease the performance of network inference algorithms
  For the \ac{op} step, we recommend that the filtering be turned on and used at the default settings, unless a taxonomy entity of interest might be filtered out at the default settings.
  This is necessary since a large number of rows in the counts matrix would reduce the accuracy of the inference methods and increase the computational time.
  Finally for the \ac{ni}, we recommend the scaled-sum consensus method along with p-value merging (for the correlation-based methods) to generate a final consensus network for the input data.

### 2022-04-25 sections/supplementary.tex
Add: Add captions and update supplementary figures

    % TODO: Elaborate on this
    Figure~\ref{fig:figure_s1} describes the variation in the inferrred networks in the first two principal axes.
    This Figure is the similar to Figure~\ref{fig:figure2}B and further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.

    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s1.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The variation in the inferred networks in the first two principal axes}
          The plots are colored by the tools or parameters used in the \ac{dc} step (A) and the \ac{op} step (B)
        }
      \label{fig:figure_s1}
    \end{figure}
    \FloatBarrier
    \newpage

    % TODO: Write something about TSNE
    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s2.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The TSNE plot of the networks inferred by various methods}.
        }
      \label{fig:figure_s2}
    \end{figure}
    \FloatBarrier
    \newpage

  % QUESTION: Should the CCA section be only included in the response letter?

@@

    Figure~\ref{fig:figureS3} is similar to Figure~\ref{fig:figure3} shows the average UniFrac distance between reference sequences generated by the various methods in the \ac{dc} step.
    In Figure~\ref{fig:figureS3} however, we use only the top 1000 representative sequences (by abundance) for this calculation.
    We observe that both the weighted and unweighted UniFrac distances are increased compared to the figure with the all the representative sequences.
    This shows that the most abundant representative sequences generated by the methods are not similar to each other.

    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s3.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The UniFrac distance between the 100 most abundant representative sequences generated by various method in the \ac{dc} step}
        }
      \label{fig:figure_s3}
    \end{figure}
    \FloatBarrier
    \newpage

    We find that the two chimera checking methods, uchime and remove bimera, produce similar outputs.
    This can be verified by comparing the output (representative sequence distributions) of the two chimera checking methods for each denoising method (Figure~\ref{fig:figure_s4}).
    We find that with the exception of de novo and open-reference under the unweighted UniFrac metric, all the other methods have low dissimilarity.
    This is especially true for the \ac{dada2} and Deblur methods which are the recommended denoising methods in the \ac{micone} pipeline.
    % TODO: Can we verify this?
    Therefore, we recommended remove bimera as the default chimera method if one is using \ac{dada2} and uchime when one is using Deblur, since these methods were developed for these respective algorithms.

    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s4.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
        }
      \label{fig:figure_s4}
    \end{figure}
    \FloatBarrier
    \newpage

@@

    % TODO: Elaborate on this and work on the caption
    Figure~\ref{fig:figure_s5} is similar to Figure~\ref{fig:figure4}B, but instead of the matching only the top 100 taxonomic entities (by abundance), we match all the assignments from one database with those from the other two databases.
    We observe that the percentage of mismatches are indeed higher when considering all the assignments, implying that matching of the taxonomies in the more abundant sequences are more consistent.

    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s5.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
        }
      \label{fig:figure_s5}
    \end{figure}
    \FloatBarrier
    \newpage

@@

  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|}

@@

    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s6.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
        }
      \label{fig:figure_s6}
    \end{figure}
    \FloatBarrier
    \newpage

@@

    % TODO: Elaborate on this
    In Figure~\ref{fig:figure_s7} we perform the same analysis we did with the NorTA data using the seqtime data.
    We observe that the best performing network inference algorithm is \ac{spieceasi}, but the consensus algorithms, both the scaled-sum and simple voting methods outperform it in terms of precision.
    These results show that the scaled-sum method is much better suited for inferring robust and accurate interactions from count data regardless of the distributions, topologies and number of samples in the data.

    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s7.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
        }
      \label{fig:figure_s7}
    \end{figure}
    \FloatBarrier
    \newpage

>>>>

    Figure~\ref{fig:figure_s1} describes the variation in the inferred networks in the first two principal axes and is similar to Figure~\ref{fig:figure2}B.
    The \ac{dc} step does not seem to have any correlation with the similarity of networks on the PCA plot, but all the networks with filtering at the \ac{op} step are similar to each other.
    This further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.

    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s1.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The variation in the inferred networks in the first two principal axes}
          All combinations of inferred networks are shown as points on a PCA plot.
          Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
          The points are colored by the tools or parameters used in the \ac{dc} step (A) and the \ac{op} step (B).
        }
      \label{fig:figure_s1}
    \end{figure}
    \FloatBarrier
    \newpage

    % TODO: Verify this about t-SNE
    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s2.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The t-SNE plot of the networks inferred by various methods}.
          The points are colored by the tools and parameters used in \ac{dc} step (A), \ac{ta} step (B), \ac{op} step (C) and \ac{ni} step (D).
          The separation of the points based on taxonomy reference database shows that the points might cluster based on reference database in high-dimensional space.
        }
      \label{fig:figure_s2}
    \end{figure}
    \FloatBarrier
    \newpage

@@

    Figure~\ref{fig:figure_s3} is similar to Figure~\ref{fig:figure3} shows the average UniFrac distance between reference sequences generated by the various methods in the \ac{dc} step.
    In Figure~\ref{fig:figure_s3} however, we use only the top 1000 representative sequences (by abundance) for this calculation.
    We observe that both the weighted and unweighted UniFrac distances are increased compared to the figure with the all the representative sequences.
    This shows that the 1000 most abundant representative sequences generated by the methods are not similar to each other.
    And since the weighted UniFrac is much smaller than the unweighted UniFrac distance, we can conclude that those reference sequences that are present in the middle of the abundance distribution are dissimilar.

    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s3.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The UniFrac distance between the 1000 most abundant representative sequences generated by various method in the \ac{dc} step}
        }
      \label{fig:figure_s3}
    \end{figure}
    \FloatBarrier
    \newpage

    We find that the two chimera checking methods, uchime and remove bimera, produce similar outputs.
    This can be verified by comparing the output (representative sequence distributions) of the two chimera checking methods for each denoising method (Figure~\ref{fig:figure_s4}).
    We find that with the exception of de novo and open-reference under the unweighted UniFrac metric, all the other methods have low dissimilarity.
    This is especially true for the \ac{dada2} and Deblur methods which are the recommended denoising methods in the \ac{micone} pipeline.
    Therefore, we recommended remove bimera as the default chimera method if one is using \ac{dada2} and uchime-denovo when one is using Deblur, since these methods were developed for these respective algorithms (\ac{qiime2} uses uchime-denovo in the Deblur workflow).

    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s4.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The weighted and unweighted UniFrac distance between representative sequences generated using remove bimera and uchime for each denoising method.}
        }
      \label{fig:figure_s4}
    \end{figure}
    \FloatBarrier
    \newpage

@@

    Figure~\ref{fig:figure_s5} is similar to Figure~\ref{fig:figure4}B, but instead of the matching only the top 100 taxonomic entities (by abundance), we match all the assignments from one database with those from the other two databases.
    We observe that the percentage of mismatches are indeed higher when considering all the assignments, implying that matching of the taxonomies in the more abundant sequences are more consistent.

    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s5.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The pairwise comparison of assignments generated using different databases for each representative sequence}
        }
      \label{fig:figure_s5}
    \end{figure}
    \FloatBarrier
    \newpage

@@

    \centering
    \small
    \begin{tabular}{|l|l|l|l|l|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}

@@

    % TODO: Rerun this figure
    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s6.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The precision and sensitivity of the inferred networks on the NorTA synthetic interaction data with Pearson and Spearman added to the consensus.}
            The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
            The different consensus based methods used are: pvalue merging, scaled-sum (SS) and simple voting (SV) method.
            With the addition of Pearson and Spearman in the consensus calculation we observe a decrease in the average precision scores of both the consensus algorithms.
            The overall best precision was still consistently obtained by the scaled-sum method (0.985, 0.963 and 1.000), but the performance was not stable to changes in parameter value.
            The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), again outperforms \ac{spieceasi} on average precision (0.961).
        }
      \label{fig:figure_s6}
    \end{figure}
    \FloatBarrier
    \newpage

@@

    In Figure~\ref{fig:figure_s7} we perform the same analysis we did with the NorTA data using the seqtime data.
    We observe that the best performing network inference algorithm is \ac{spieceasi}, but the consensus algorithms, both the scaled-sum and simple voting methods outperform it in terms of precision.
    These results show that the scaled-sum method is much better suited for inferring robust and accurate interactions from count data regardless of the distributions, topologies and number of samples in the data.

    \begin{figure}[H]
      \centering
      \includegraphics[width=1.0\linewidth]{figure_s7.pdf}
    \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The precision and sensitivity of the inferred networks on the seqtime synthetic interaction data.}
          The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
          The different consensus based methods used are: pvalue merging, scaled-sum (SS) and simple voting (SV) method.
          Pearson and Spearman methods are not used in the calculation of the consensus.
          Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.624), but the overall best precision was consistently obtained by the scaled-sum method (0.688, 0.820 and 1.000).
          The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), also outperforms \ac{spieceasi} on average precision (0.692).
        }
      \label{fig:figure_s7}
    \end{figure}
    \FloatBarrier
    \newpage
